---
title: "Trips by county"
author: "Sunaina Martin"
date: "11/8/2021"
output: html_document
---
####Load the data into R

This dataset from the Department of Transport Statistics uses cellular data to track people who spend more than 10 mins away from a home location.

```{r}
library(tidyverse)
trips_by_county<- read.csv('https://raw.githubusercontent.com/hkular/stats_project/main/trips_by_county.csv')
view(trips_by_county)
summary(trips_by_county)
```

####change date from characters to date format

```{r}
trips_by_county$Date<-as.Date(trips_by_county$Date) 
```

####Rename columns for easy access

```{r}
trips_by_county<-trips_by_county%>%
  rename( postal_code = State.Postal.Code,
        county_name = County.Name,
        population_home = Population.Staying.at.Home,
        pop_not_home = Population.Not.Staying.at.Home
      
                        )
```

####Bin trips into short, medium or long to make it things easier. 

Short trips could reflect people taking walks, going to the grocery, biking, going to work etc. Med trips are likely work-related or travel. Long trips are also likely travel to see friends and family or in some cases for work.

```{r}
trips_by_county$short_trips<- (trips_by_county$Number.of.Trips..1 + trips_by_county$Number.of.Trips.1.3 + trips_by_county$Number.of.Trips.3.5 + trips_by_county$Number.of.Trips.5.10)/1000
trips_by_county$med_trips<-(trips_by_county$Number.of.Trips.10.25+trips_by_county$Number.of.Trips.25.50)/1000
trips_by_county$long_trips<- (trips_by_county$Number.of.Trips.50.100 + trips_by_county$Number.of.Trips.250.500 + trips_by_county$Number.of.Trips...500)/1000
view(trips_by_county)

```
 
####Load movement data for 2019 for comparison and perform the same manipulations on this dataframe as above. 

```{r}
trips_by_county_2019<- read.csv('https://raw.githubusercontent.com/hkular/stats_project/main/Trips_by_Distance%202019.csv')
view(trips_by_county_2019)
summary(trips_by_county_2019)
trips_by_county_2019$Date<-as.Date(trips_by_county_2019$Date) 
trips_by_county_2019<-trips_by_county_2019%>%
  rename( postal_code = State.Postal.Code,
        county_name = County.Name,
        population_home = Population.Staying.at.Home,
        pop_not_home = Population.Not.Staying.at.Home)
trips_by_county_2019$short_trips<- (trips_by_county_2019$Number.of.Trips..1 + trips_by_county_2019$Number.of.Trips.1.3 + trips_by_county_2019$Number.of.Trips.3.5 + trips_by_county_2019$Number.of.Trips.5.10)/1000
trips_by_county_2019$med_trips<-(trips_by_county_2019$Number.of.Trips.10.25+trips_by_county_2019$Number.of.Trips.25.50)/1000
trips_by_county_2019$long_trips<- (trips_by_county_2019$Number.of.Trips.50.100 + trips_by_county_2019$Number.of.Trips.250.500 + trips_by_county_2019$Number.of.Trips...500)/1000
view(trips_by_county_2019)


``` 

####Number of short trips taken as a function of location pre-pandemic

```{r}
short_trips_by_county_2019<-  ggplot(data = trips_by_county_2019, aes(x=Date, y = short_trips, fill = county_name))  +
  geom_smooth() + 
  labs(x = "Date", y = "Number of trips taken (in the thousands)", title = "Short trips taken as a function of the county pre-pandemic", fill = "County") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
                                                                                                        
short_trips_by_county_2019
```


####Number of medium trips taken as a function of location pre-pandemic

```{r}
med_trips_by_county_2019<-  ggplot(data = trips_by_county_2019, aes(x=Date, y = med_trips, fill = county_name))  +
  geom_smooth() + 
  labs(x = "Date", y = "Number of trips taken (in the thousands)", title = "Mid-distance trips taken as a function of the county pre-pandemic", fill = "County") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
                                                                                                        
med_trips_by_county_2019

```

####Number of long distance trips taken as a function of location pre-pandemic

```{r}
long_trips_by_county_2019<-  ggplot(data = trips_by_county_2019, aes(x=Date, y = long_trips, fill = county_name))  +
  geom_smooth() + 
  labs(x = "Date", y = "Number of trips taken (in the thousands)", title = "Long distance trips taken as a function of the county pre-pandemic", fill = "County") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
                                                                                                        
long_trips_by_county_2019
```

####Number of short trips taken as a function of location during the pandemic

```{r}
short_trips_by_county_2020<-  ggplot(data = trips_by_county, aes(x=Date, y = short_trips, fill = county_name))  +
  geom_smooth() + 
  labs(x = "Date", y = "Number of trips taken (in the thousands)", title = "Short trips taken as a function of the county during pandemic", fill = "County") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
                                                                                                        
short_trips_by_county_2020
```


####Number of medium trips taken as a function of location during the pandemic

```{r}
med_trips_by_county_2020<-  ggplot(data = trips_by_county, aes(x=Date, y = med_trips, fill = county_name))  +
  geom_smooth() + 
  labs(x = "Date", y = "Number of trips taken (in the thousands)", title = "Mid-distance trips taken as a function of the county during pandemic", fill = "County") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
                                                                                                        
med_trips_by_county_2020
```

####Number of long distance trips taken as a function of location during the pandemic

```{r}
long_trips_by_county_2020<-  ggplot(data = trips_by_county, aes(x=Date, y = long_trips, fill = county_name))  +
  geom_smooth() + 
  labs(x = "Date", y = "Number of trips taken (in the thousands)", title = "Long distance trips taken as a function of the county during pandemic", fill = "County") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
                                                                                                        
long_trips_by_county_2020
```

